Procurements of the Canadian Government, Runaway Military Contracts

### 2024-12-03 paper/paper.qmd
added custom colors + fixed first 3 graphs

The head of the IFJ Database:

```{r fig-head1}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

df_table1 <- df %>%
  select(
    Contract = contract,
    Buyer = buyer,
    Supplier = supplier,
    Amount = amount,
    `Award Date` = award_date,
    `Start Date` = start_date,
    `End Date` = end_date
  )

@@

# Display the first table
kable(head(df_table1_truncated, 5), caption = "Sample of the Data - Part 1") %>%
  kable_styling(
    font_size = 10,
    latex_options = c("scale_down")
  )

@@

```{r fig-head2}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

df_table2 <- df %>%
  select(
    `Baseline Days` = baseline_days,
    `Duration Days` = duration_days,
    `Processed Supplier` = processed_supplier
  )

@@

# Display the second table
kable(head(df_table2_truncated, 5), caption = "Sample of the Data - Part 2") %>%
  kable_styling(
    font_size = 10,
    latex_options = c("scale_down")
  )

@@

```{r fig-pie-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Top 5 Buyers by Total Amount Spent and Others (Pie Chart)
#| fig-align: center

@@

# TODO: Assign specific colors to each buyer group
# Example color assignments (replace with desired colors)
buyer_colors <- c(
  "National Defence" = "#1f77b4",
  "Public Services and Procurement Canada" = "#ff7f0e",
  "Shared Services Canada" = "#2ca02c",
  "Public Health Agency of Canada" = "#d62728",
  "Fisheries and Oceans Canada" = "#9467bd",
  "Other"  = "#8c564b"
)

# Plot the pie chart
ggplot(buyer_totals_grouped, aes(x = "", y = total_amount, fill = buyer_group)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  labs(
    title = "Top 5 Buyers by Total Amount Spent and Others",
    fill = "Buyer"
  ) +
  scale_fill_manual(values = buyer_colors) +
  theme_void() + # Removes background, grid, and numeric labels
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 4)
```

@@

```{r fig-mil-spending}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

@@

# Summarize total spending per year
military_spending_per_year <- military_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE))

# Create the bar chart
ggplot(data = military_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Military Spending Per Year", x = "Year", y = "Total Spending") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

Public Services and Procurements Canada is the clearinghouse for the entire federal government. This means they are responsible for managing the procurement process for buyers across the government [REF]. It is important to note that the PSPC sometimes does not publish small contracts, less than /$40,000.  PSPC spending over time shows while military spending increases the rest of the government is taking a hit.
```{r fig-public-spending}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

@@

# Summarize total spending per year
public_spending_per_year <- public_services_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE))

# Create the bar chart
ggplot(data = public_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Public Services Spending Per Year", x = "Year", y = "Total Spending") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r fig-top-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center

@@

# Optional: Display the summarized data
print(total_spent_df)

@@

RSelenuim
VGAM
stringr
dplyr
tidyverse
zoo
kableExtra
knitr
scales
lubridate
ggplot2
rvest
readr
testthat

>>>>

The head of the IFJ Database:
```{r fig-head1}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Sample of the Data
#| fig-align: center
#| fig-pos: "H"

df_table1 <- df %>%
  select(
    Contract = contract,
    Buyer = buyer,
    Supplier = supplier,
    Amount = amount,
    `Award Date` = award_date,
    `Start Date` = start_date,
    `End Date` = end_date
  ) %>%
  mutate(
    Amount = comma(Amount)  # Format Amount with commas
  )

@@

# Display the first table
kable(head(df_table1_truncated, 5)) %>%
  kable_styling(
    font_size = 10,
    latex_options = c("scale_down")
  )

@@

```{r fig-head2}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Sample of Derived Data
#| fig-align: center
#| fig-pos: "H"

df_table2 <- df %>%
  select(
    `Baseline Days` = baseline_days,
    `Duration Days` = duration_days,
    `Processed Supplier` = processed_supplier
  ) %>%
  mutate(
    `Baseline Days` = comma(`Baseline Days`),  # Format Baseline Days with commas
    `Duration Days` = comma(`Duration Days`)   # Format Duration Days with commas
  )

@@

# Display the second table
kable(head(df_table2_truncated, 5)) %>%
  kable_styling(
    font_size = 10,
    latex_options = c("scale_down")
  )

@@

```{r fig-pie-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Top 5 Buyers by Total Amount Spent and Others
#| fig-align: center
#| fig-pos: "H"

colors <- c("#377eb8", "#e41a1c", "#4daf4a", "#ff7f00", "#984ea3", "#808080")

@@

# TODO: Assign specific colors to each buyer group
# Example color assignments (replace with desired colors)
buyer_colors <- c(
  "National Defence" = "#e41a1c",
  "Public Services and Procurement Canada" = "#984ea3",
  "Shared Services Canada" = "#ff7f00",
  "Public Health Agency of Canada" = "#4daf4a",
  "Fisheries and Oceans Canada" = "#377eb8",
  "Other"  = "#808080"
)

# Plot the pie chart
ggplot(buyer_totals_grouped, aes(x = "", y = total_amount, fill = buyer_group)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  labs(
    fill = "Buyer"
  ) +
  scale_fill_manual(values = buyer_colors) +
  theme_void() + # Removes background, grid, and numeric labels
  theme(
    legend.position = "right",
    plot.title = element_text(hjust = 0.5, size = 16, face = "bold")
  ) +
  geom_text(aes(label = paste0(round(percentage, 1), "%")),
            position = position_stack(vjust = 0.5),
            color = "white",
            size = 2.5)
```

@@

```{r fig-mil-spending}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Military Spending Per Year in Billions
#| fig-align: center
#| fig-pos: "H"

@@

military_spending_per_year <- military_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE) / 1e9)

# Create the bar chart
ggplot(data = military_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "#e41a1c") +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

Public Services and Procurements Canada is the clearinghouse for the entire federal government. This means they are responsible for managing the procurement process for buyers across the government [REF]. It is important to note that the PSPC sometimes does not publish small contracts, less than /$40,000.  PSPC spending over time shows while military spending increases the rest of the government is taking a hit.
```{r fig-public-spending}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: Public Services Spending Per Year in Billions
#| fig-align: center
#| fig-pos: "H"

@@

# Summarize total spending per year
public_spending_per_year <- public_services_df %>%
  group_by(year) %>%
  summarize(total_spending = sum(amount, na.rm = TRUE)  / 1e9 )

# Create the bar chart
ggplot(data = public_spending_per_year, aes(x = year, y = total_spending)) +
  geom_bar(stat = "identity", fill = "#984ea3") +
  labs(x = "Year", y = "Total Spending (in Billions)") +
  theme_minimal() +
  scale_y_continuous(labels = dollar_format(prefix = "$")) +
  theme(plot.title = element_text(hjust = 0.5))

@@

```{r fig-top-buyers}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

@@

```{r fig-top-supps}
#| message: false
#| echo: false
#| warning: false
#| fig-cap: template
#| fig-align: center
#| fig-pos: "H"

@@

@@

RSelenuim
VGAM
stringr
dplyr
tidyverse
zoo
kableExtra
knitr
scales
lubridate
ggplot2
rvest
readr
testthat
colorbrewerV2
